---
title: "R, Group A: Practical work"
format: html
editor: visual
---

## Exercise 1

The EXCEL workbook SelectoStudy.xlsx contains several variables of a cross-sectional study, the so-called "Selecto Study", on the association between cumulative ecstasy use and cognitive dysfunction, which was car- ried from 2008 to 2010 in the *Hospital del Mar Research Institute* (<https://www.imim.es/en_index.html)1.> The workbook has three sheets that contain the data of each of the three study groups (Ecstasy (XTC)) users, cannabis users, and non-users (control group)) and a fourth sheet with the labels of the variables and some additional information on the neuropsychological tests the study participants had to do.

### (a)

Import the data of the three study groups into R and create a single data frame with name selecto that contains all data and also a variable (group) that identifies the study group of each volunteer. This variable should be a factor whose first level corresponds to the group of ecstasy users. In addition, variables sex and city should also be factors.

```{r}
library(openxlsx)
library(dplyr)
library(lubridate)
library(descr)
```

```{r}
XTC <- read.xlsx("SelectoStudy.xlsx",sheet = "XTC",startRow = 3)
Cannabis <- read.xlsx("SelectoStudy.xlsx",sheet = "Cannabis",startRow = 3)
Control <- read.xlsx("SelectoStudy.xlsx",sheet = "Control",startRow = 3)

XTC <- mutate(XTC, "group" = "XTC")
Cannabis <- mutate(Cannabis, "group" = "Cannabis")
Control <- mutate(Control, "group" = "Control")

selecto <- rbind(XTC, Cannabis, Control)

selecto[ , c("sex", "city", "group")] <- 
  lapply(selecto[ , c("sex", "city", "group")] , factor)

```

### (b)

How many volunteers are in each group?

```{r}
selecto %>%
  group_by(group) %>%
  summarise(number_of_volunteers = n())
```

### (c)

How many volunteers with compete information, i.e., no missing values, are in each group?

```{r}
na.omit(selecto) %>%
  group_by(group) %>%
  summarise(number_of_volunteers = n())

```

### (d)

Which volunteer has most missing data?

```{r}
selecto[which.max(rowSums(is.na(selecto))), ]
```

### (e)

Convert the variables birth and studydate in date variables and add the volunteers' ages as a new variable to the data frame.

```{r}
selecto <- selecto %>%
  mutate(
    birth = as.Date(birth, format = "%Y-%m-%d"),
    studydat = as.Date(studydat, format = "%Y-%m-%d"),
    age = as.vector(trunc(difftime(studydat, birth) / 365.25))
  )

```

### (f)

Order the rows of the data frame by study group, gender, and decreasing age.

```{r}
selecto <- selecto %>%
  arrange(group, sex, desc(age))
```

### (g)

Which are the ages of the first and last volunteer and from which cities are they?

```{r}
selecto[c(1, nrow(selecto)), c("age", "city")]
```

### (h)

Choose randomly two volunteers in each group and replace their residence by Gavà

```{r}
set.seed(123)

selecto$city <- as.character(selecto$city)
selecto[sample(nrow(selecto), 2), "city"] <- "Gava"

```

### (i)

Which is the proportion of volunteers in each group that were born on a weekend?

```{r}
Sys.setlocale("LC_TIME", "English")
selecto %>%
  mutate(weekday = weekdays(selecto$birth)) %>%
  group_by(group) %>%
  summarize(proportion_weekend = mean(weekday %in% c("Saturday", "Sunday"), 
                                      na.rm = TRUE))
```

### (j)

Create the ordinal variable agecat that groups the volunteers' ages in four categories: ≤ 20, 21 -- 25, 26 -- 30, and \> 30 years.

```{r}
agecat_fun <- function(age) {
  if (age <= 20) {
    return("young")
  }
  if ((age > 20) & (age < 26)){
    return("young adult")
  }
  if ((age > 25) & (age < 31)){
    return("adult")
  }
  else{
    return("old")
  }
}

selecto <- selecto %>%
  mutate(agecat = sapply(age, agecat_fun))
```

### (k)

Use the *CrossTable* function of the descr package to represent the conditional distribution of the variable agecat per study group

```{r}
CrossTable(selecto$group,selecto$agecat)
```

### (l)

Write the data frame selecto to a new sheet named \'Selecto\' of the EXCEL workbook Selecto.xlsx.

```{r}
write.xlsx(selecto,"Selecto.xlsx",sheetName = "Selecto",append = FALSE)
```

### (m)

Program a pipe with the operator \|\> to create the following subset of the data frame selecto and write it
to a new sheet named \'SubsetSelecto\' of the EXCEL workbook Selecto.xlsx. The subset should contain
the data of those volunteers in the ecstasy or cannabis group that were born in the 80ies and whose
switS1 score was known. In addition, the scores of the neuropsychological variables cvltot, reymi,
reymd, reycop, and animtotal should be scaled so that they have mean 0 and standard deviation 1, and
the variables birth and studydat should be removed.

```{r}
SubsetSelecto <- selecto %>% filter(group != "Control") %>% 
  filter(between(birth,as.Date('1980-01-01'),as.Date('1989-12-31'))) %>%
  filter(!is.na(switS1)) %>%
  mutate(birth = NULL, studydat = NULL,cvltot = scale(cvltot),
         reymd = scale(reymd),reycop = scale(reycop),animtotal = scale(animtotal))
{r}
wb <- createWorkbook()
addWorksheet(wb, "Selecto")
addWorksheet(wb, "SubsetSelecto")

writeData(wb, "Selecto",selecto, startRow = 1, startCol = 1)
writeData(wb, "SubsetSelecto",SubsetSelecto, startRow = 1, startCol = 1)

saveWorkbook(wb, file = "Selecto.xlsx", overwrite = TRUE)
```

## Exercise 2
