---
title: "R, Group A: Practical work"
format: html
editor: visual
---

## Exercise 1

The EXCEL workbook SelectoStudy.xlsx contains several variables of a cross-sectional study, the so-called "Selecto Study", on the association between cumulative ecstasy use and cognitive dysfunction, which was car- ried from 2008 to 2010 in the *Hospital del Mar Research Institute* (<https://www.imim.es/en_index.html)1.> The workbook has three sheets that contain the data of each of the three study groups (Ecstasy (XTC)) users, cannabis users, and non-users (control group)) and a fourth sheet with the labels of the variables and some additional information on the neuropsychological tests the study participants had to do

### (a)

Import the data of the three study groups into R and create a single data frame with name selecto that contains all data and also a variable (group) that identifies the study group of each volunteer. This variable should be a factor whose first level corresponds to the group of ecstasy users. In addition, variables sex and city should also be factors.

```{r}
library(openxlsx)
library(dplyr)
library(lubridate)


```

```{r}
XTC <- read.xlsx("SelectoStudy.xlsx",sheet = "XTC",startRow = 3)
Cannabis <- read.xlsx("SelectoStudy.xlsx",sheet = "Cannabis",startRow = 3)
Control <- read.xlsx("SelectoStudy.xlsx",sheet = "Control",startRow = 3)

XTC <- mutate(XTC, "group" = "XTC")
Cannabis <- mutate(Cannabis, "group" = "Cannabis")
Control <- mutate(Control, "group" = "Control")

selecto <- rbind(XTC, Cannabis, Control)

selecto[ , c("sex", "city", "group")] <- 
  lapply(selecto[ , c("sex", "city", "group")] , factor)

```

### (b)

How many volunteers are in each group?

```{selecto %>%}
  group_by(group) %>%
  summarise(number_of_volunteers = n())
```

### (c)

How many volunteers with compete information, i.e., no missing values, are in each group?

```{r}
na.omit(selecto) %>%
  group_by(group) %>%
  summarise(number_of_volunteers = n())

```

### (d)

Which volunteer has most missing data?

```{r}
selecto[which.max(rowSums(is.na(selecto))), ]
```

### (e)

Convert the variables birth and studydate in date variables and add the volunteers' ages as a new variable to the data frame.

```{r}
selecto <- selecto %>%
  mutate(
    birth = as.Date(birth, format = "%Y-%m-%d"),
    studydat = as.Date(studydat, format = "%Y-%m-%d"),
    age = as.vector(trunc(difftime(studydat, birth) / 365.25))
  )

```

### (f)

Order the rows of the data frame by study group, gender, and decreasing age.

```{r}
selecto <- selecto %>%
  arrange(group, sex, desc(age))
```

### (g)

Which are the ages of the first and last volunteer and from which cities are they?

```{r}
selecto[c(1, nrow(selecto)), c("age", "city")]
```

### (h)

Choose randomly two volunteers in each group and replace their residence by Gav\`a

```{r}
set.seed(123)

selecto$city <- as.character(selecto$city)
selecto[sample(nrow(selecto), 2), "city"] <- "Gava"

```

### (i)

Which is the proportion of volunteers in each group that were born on a weekend?

```{r}
selecto %>%
  mutate(weekday = weekdays(selecto$birth)) %>%
  group_by(group) %>%
  summarize(proportion_weekend = mean(weekday %in% c("Saturday", "Sunday"), 
                                      na.rm = TRUE))
```

### (j)

Create the ordinal variable agecat that groups the volunteers' ages in four categories: ≤ 20, 21 -- 25, 26 -- 30, and \> 30 years.

```{r}
agecat_fun <- function(age) {
  if (age <= 20) {
    return("young")
  }
  if ((age > 20) & (age < 26)){
    return("young adult")
  }
  if ((age > 25) & (age < 31)){
    return("adult")
  }
  else{
    return("old")
  }
}

selecto <- selecto %>%
  mutate(agecat = sapply(age, agecat_fun))
```
